---
title: "Analyzing Forest Fire Data"
author: "Ana Alvarez"
date: "21/1/2021"
output: html_document
---
```{r}
library(tidyverse)

# Exploring data
forest_fires <- read_csv("forestfires.csv")

forest_fires %>% 
  head()
```
The data we'll be working with is associated with a scientific research paper on predicting the occurrence of forest fires in Portugal using modeling techniques. 
Source: http://www3.dsi.uminho.pt/pcortez/fires.pdf

We'll perform exploratory analyses on the data to better understand it and any relationships that might be present in it.

The variables in the dataset are:

* X: X-axis spatial coordinate within the Montesinho park map: 1 to 9
* Y: Y-axis spatial coordinate within the Montesinho park map: 2 to 9
* month: Month of the year: 'jan' to 'dec'
* day: Day of the week: 'mon' to 'sun'
* FFMC: Fine Fuel Moisture Code index from the FWI system: 18.7 to 96.20
* DMC: Duff Moisture Code index from the FWI system: 1.1 to 291.3
* DC: Drought Code index from the FWI system: 7.9 to 860.6
* ISI: Initial Spread Index from the FWI system: 0.0 to 56.10
* temp: Temperature in Celsius degrees: 2.2 to 33.30
* RH: Relative humidity in percentage: 15.0 to 100
* wind: Wind speed in km/h: 0.40 to 9.40
* rain: Outside rain in mm/m2 : 0.0 to 6.4
* area: The burned area of the forest (in ha): 0.00 to 1090.84

**FWI**: "fire weather index", a method used by scientists to quantify risk factors for forest fires.

```{r}
# exploring the month and day variable
forest_fires %>% 
  pull(month) %>% 
  unique

forest_fires %>% 
  pull(day) %>% 
  unique

# convert month and day to categorical variables with an order

month_levels = c("jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")
day_levels = c("mon", "tue", "wed", "thu", "fri", "sat", "sun")

forest_fires <- forest_fires %>% 
  mutate(
    month = factor(month, levels = month_levels),
    day = factor(day, levels = day_levels)
  )

```
# Which months do forest fires happen the most? 
```{r}
fires_per_month <- forest_fires %>% 
  group_by(month) %>% 
  summarise(n_fires = n())

fires_per_day <- forest_fires %>% 
  group_by(day) %>% 
  summarise(n_fires = n())

fires_per_month

fires_per_month %>% 
  ggplot(aes(x = month, y = n_fires, fill = -n_fires)) +
  geom_col() +
  labs(title = "Number of fires per month",
       y = "Number of fires"
       ) +
  scale_fill_continuous(name = "Number of fires")

```
August and September are the months with the highest number of fires, by far.

```{r}
fires_per_day %>% 
  ggplot(aes(x= day, y = n_fires, fill = -n_fires)) +
  geom_col() +
  labs(title = "Number of fires per day ",
       y = "Number of fires") +
  scale_fill_continuous(name = "Number of fires")
```
Sunday, Friday and Saturday are the days of the week with more probabilities of fires. That points to human responsibility of those fires.

# Relationships between the variables and time (months)

```{r}
colnames(forest_fires)

forest_fires_long <- forest_fires %>% 
  pivot_longer(cols = c("FFMC", "DMC", "DC", "ISI", "temp", "RH", "wind", "rain"),
               names_to = "conditions",
               values_to = "value")

forest_fires_long %>% 
  ggplot(aes(x = month, y = value)) +
  geom_boxplot() +
  facet_wrap(vars(conditions),
             scales = "free_y") +
  theme_light()
  
```
We'll use this variable area as an indicator of the severity of the fire. 

# Relationships between the area burned by a forest fire and the other variables

```{r}
forest_fires_long %>% 
  ggplot(aes(x = value, y = area)) +
  geom_point() +
  facet_wrap(vars(conditions),
             scales = "free_x") +
  labs(title = "Relationship between environmental conditions and severity of fire",
       ylab = "Area burnes (hectare)") +
  theme_light()

```
Absence of rain has a direct relationship with severity of fire. Also high temperature, low humidity and high Fine Fuel Moisture Code.

```{r}
forest_fires_long %>% 
  filter(area < 300) %>% 
  ggplot(aes(x = value, y = area)) +
  geom_point() +
  facet_wrap(vars(conditions),
             scales = "free_x") +
  labs(title = "Relationship between environmental conditions and severity of fire",
       ylab = "Area burnes (hectare)") +
  theme_light()
```
If we remove outliers, setting a limit of 300 hectare for area extension, we can see direct relationship between absence of rain and quantity of fires. But no clear relationships for the other variables. 


