---
title: "Book Sales"
author: "Ana Alvarez"
date: "20/1/2021"
output: html_document
---
```{r}
library(tidyverse)
library(lubridate)


# Exploring data

sales_2019 <- read_csv("sales2019.csv")

dim(sales_2019)

colnames(sales_2019)

for (col in colnames(sales_2019)) {
  paste0(col, " : ", typeof(sales_2019[[col]])) %>% 
    print
}

glimpse(sales_2019)

sales_2019 %>% 
  filter(is.na(title))

```
The dataset has `5000` rows and `5` columns. All columnas are character type except the `total_purchased` one, that's numerical. Date column has a character type. `total_purchased` and `user_submitted_review` has missing data.

## Remove missing data in user_submitted_review
```{r}

# `885` rows have missing values in the `user_submitted_review`. We store them before cleaning the dataset.
na_user_review <- sales_2019 %>% 
  filter(is.na(user_submitted_review))

sales_2019_clean <- sales_2019 %>% 
  filter(!(is.na(user_submitted_review)))
```

## Average number of books sold per order
```{r}
avg_order <- sales_2019_clean %>% 
  summarise(avg_order = mean(total_purchased, na.rm = TRUE))

avg_order <- avg_order[[1]]

avg_order
```
## Assign average order to missing values in total_purchased
```{r}
sales_2019_clean <- sales_2019_clean %>% 
  mutate(
    imputed_purchases = if_else(is.na(total_purchased), avg_order, total_purchased)
    )

sales_2019_clean
```

## Classify reviews as either positive or negative
```{r}
# unique sentences in `user_submitted_reviews`
unique(sales_2019_clean$user_submitted_review)

sales_2019_clean <- sales_2019_clean %>% 
  mutate(classification = case_when(
    str_detect(user_submitted_review, "okay") ~ "possitive",
    str_detect(user_submitted_review, "Awesome") ~ "possitive",
    str_detect(user_submitted_review, "OK") ~ "possitive",
    str_detect(user_submitted_review, "learned") ~ "possitive",
    str_detect(user_submitted_review, "Never") ~ "possitive",
    TRUE ~ "negative"
  ))

sales_2019_clean %>% 
  count(classification) %>% 
  arrange(-n)

```
## Was the new book program effective in increasing book sales?
The program started on July 1st, 2019. We want to analyze is sales increase after this date.

```{r}
# conversion of the date column

sales_2019_clean

sales_2019_clean <- sales_2019_clean %>% 
  mutate(date = mdy(date))
  
```
```{r}
# Column to show if purchase is after or before the program started

program_date <- ymd("20190701")

comprobation_date <- function(a_date){
  if(a_date >= program_date){return("after")
    }else{return("before")}
}

sales_2019_clean <- sales_2019_clean %>% 
  mutate(program = unlist(map(date, comprobation_date)))

sales_2019_clean

sales_2019_clean %>% 
  group_by(program) %>% 
  summarise(amount = sum(imputed_purchases))

sales_2019_clean %>% 
  group_by(program, customer_type) %>% 
  summarise(total_sales = sum(imputed_purchases))

sales_2019_clean %>% 
  group_by(program, classification) %>% 
  summarise(amount = n())
```
## Conclussion

* The amount of books sold is similar before and after the program started. 
* Sales by business increased after the program. Sales by individuals decreased.
* The negative reviews increased after the beginning of the program.


# Extra analysis
Is there any relationship between month and the amount of books that were sold? 

```{r}

sales_2019_clean %>% 
  mutate(month = month(date)) %>% 
  group_by(month) %>% 
  summarise(amount = sum(imputed_purchases))

```
December was the best month in books sold of the year (Christmas gifts?). 


