---
title: "SQL Business Analysis Project"
output: html_document
---
```{r}
library(tidyverse)
library(RSQLite)
library(DBI)
```

# Create a helper function
```{r}

db <- 'chinook.db'

run_query <- function(query){
  conn <-  dbConnect(SQLite(), db)
  result <- dbGetQuery(conn, query)
  dbDisconnect(conn)
  return(result)
}

show_tables <- function(){
  q <- 'SELECT name, type FROM sqlite_master WHERE type IN ("table","view");'
  return(run_query(q))
}

show_tables()
```

# Analyzing the most sold genres in USA
```{r}
q <- ' 
WITH usa AS 
(SELECT * FROM customer WHERE country = "USA") 
SELECT 
  g.name, 
  SUM(il.quantity) n_sold, 
  CAST(SUM(il.quantity) AS FLOAT) / (select SUM(quantity) from invoice_line) pct_sold  
FROM usa  
INNER JOIN invoice i ON i.customer_id = usa.customer_id 
INNER JOIN invoice_line il ON il.invoice_id = i.invoice_id 
INNER JOIN track t ON t.track_id = il.track_id 
INNER JOIN genre g ON g.genre_id = t.genre_id 
GROUP BY g.name ORDER BY n_sold DESC;
'

genres_df <- run_query(q)

```


# Visualizing most sold genres
```{r}
genres_df %>% 
  ggplot(aes(x = reorder(name, pct_sold), y = pct_sold, fill = -pct_sold)) +
  geom_col() +
  coord_flip() +
  ylab("percentage") +
  xlab("genre") +
  theme(panel.background = element_rect(fill = "white"))
```


# Sales assigned to each sales support agent within the company

```{r}
q_sales <- 'WITH sales_agents AS 
(SELECT * FROM employee
WHERE title = "Sales Support Agent")
SELECT 
  sa.first_name || " " || sa.last_name sales_agent,
  sa.reports_to,
  sa.birthdate,
  sa.hire_date,
  sa.city,
  SUM(i.total) sales_amount
FROM sales_agents sa
INNER JOIN customer c ON c.support_rep_id = sa.employee_id
INNER JOIN invoice i ON i.customer_id = c.customer_id 
GROUP BY sa.employee_id
ORDER BY sales_amount DESC;
'

sales_per_agent <- run_query(q_sales)
sales_per_agent
```
```{r}
sales_per_agent %>% 
  ggplot(aes(x = sales_agent, y = sales_amount, fill = -sales_amount)) +
  geom_col()

```


From the information we have, the three sales agents are from the same city (Calgary), and they are hired in the same year, but the person with more experience is the agent with more sales. She is also the youngest. Female agents sale more than the male agent.

# Analyze the sales data for customers from each country
```{r}
# Total number of customers
# Total value of sales
# Average value of sales per customer
# Average order value

q_country <- 'WITH country_sales AS 
(SELECT 
  c.country,
  COUNT(distinct c.customer_id) n_customers,
  SUM(i.total) total_sales,
  SUM(i.total) / COUNT(distinct c.customer_id) avg_sale,
  COUNT(distinct i.invoice_id) n_invoices,
  SUM(i.total) / COUNT(distinct i.invoice_id) avg_invoice
FROM customer c
INNER JOIN invoice i ON i.customer_id = c.customer_id
INNER JOIN invoice_line il ON il.invoice_id = i.invoice_id
GROUP BY c.country)

SELECT 
  country_grouped,
  total_customers,
  total_sales,
  avg_sale,
  avg_invoice
FROM
(SELECT 
  CASE
    WHEN n_customers = 1 THEN "Others"
    ELSE country
    END AS country_grouped,
  CASE 
    WHEN n_customers = 1 THEN 1
    ELSE 0
    END AS sort,
  SUM(n_customers) total_customers,
  SUM(total_sales) total_sales,
  SUM(total_sales) / SUM(n_customers) avg_sale,
  SUM(total_sales) / SUM(n_invoices) avg_invoice
FROM country_sales
GROUP BY country_grouped
ORDER BY sort ASC, total_sales DESC);


'
sales_by_country <- run_query(q_country)
sales_by_country 

```


USA is the country with more total sales and more customers, followed by Canada and Brazil by far. The average sale for Czech Republic is the highest.

# Visualizing sales by country
```{r}
# Number of customers by country
sales_by_country %>% 
  ggplot(aes(x=country_grouped, y = total_customers, fill = -total_customers)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(title = "Number of customers by country", x = "Countries", y = "")

# Total sales by country
sales_by_country %>% 
  ggplot(aes(x = country_grouped, y = total_sales, fill = -total_sales)) +
  geom_col(show.legend = FALSE) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# Average sale by country
sales_by_country %>% 
  ggplot(aes(x = country_grouped, y = avg_sale, fill = -avg_sale)) +
  geom_col(show.legend = FALSE) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


```


USA is the country with most sales and Czech Republic the market with highest average sale. This two countries look the more interesting for investment.

Other small countries could look interesting for the number of sales and the total amount, by de average sale in this countries are the lowest.

# What percentage of purchases are individual tracks versus whole albums

Identifying if an invoice contains all the tracks from an album
list of tracks from an invoice and comparing it to the list of tracks from an album.

STEPS:
* Number of tracks by album.
* Group invoices by album and number of tracks of each album
* If number of tracks of each album bought equal number of tracks of the album then customer bought the whole album. Else, the customer bought individual tracks.

GOAL:
Categorize each invoice as either an album purchase or not, and calculates the following summary statistics:

*    Number of invoices
*    Percentage of invoices


```{r}
q <- '
WITH 
  songs_by_album AS
  (SELECT 
    t.album_id,
    COUNT(distinct t.track_id) n_tracks
  FROM track t 
  GROUP BY album_id),
  
  songs_by_invoice AS
(SELECT
  il.invoice_id,
  t.album_id,
  COUNT(distinct il.track_id) tracks_purchased
FROM invoice_line il
INNER JOIN track t ON t.track_id = il.track_id
GROUP BY t.album_id, il.invoice_id),

  tra_or_alb AS
(SELECT
  invoice_id,
  si.album_id,
  n_tracks,
  tracks_purchased,
  CASE WHEN
    n_tracks = tracks_purchased THEN "album"
    ELSE "tracks"
    END AS album_or_tracks
FROM songs_by_album sa
INNER JOIN songs_by_invoice si ON si.album_id = sa.album_id)

SELECT 
  album_or_tracks,
  COUNT(invoice_id) num_invoices,
  CAST(COUNT(invoice_id) AS FLOAT) / (SELECT COUNT(invoice_id) FROM tra_or_alb) pct_invoices
FROM tra_or_alb
GROUP BY album_or_tracks
'
run_query(q)
```

Most people buy tracks and not the whole album. The strategy they are considering (to purchase only the most popular tracks from each album from record companies, instead of purchasing every track from an album) looks right.







