---
title: "COVID-19 virus trends"
author: "Ana Alvarez"
date: "January 2021"
output: html_document
---
# Description
In this project we are trying to answer which countries have had the highest number of positive cases against the number of tests.

## Downloading and exploring the data

```{r}

library(readr)
library(tibble)
library(tidyverse)

covid_df <- read_csv("covid19.csv")

dim(covid_df)

vector_cols <- colnames(covid_df)
vector_cols

head(covid_df)

glimpse(covid_df)

distinct(covid_df, Country_Region) %>% 
  count()
```

The dataset contains `14` columns and `10903` rows. `109` different countries.

## Filtering data at country level

```{r}
covid_df_all_states <- covid_df %>% 
  filter(Province_State == "All States") %>% 
  select(-Province_State)
```

## Extract data from daily columns, not cummulative columns

```{r}
covid_df_all_states_daily <- covid_df_all_states %>% 
  select(Date, 
         Country_Region, 
         active, 
         hospitalizedCurr, 
         daily_tested, 
         daily_positive
         )
```

## Aggregated data by country and cases

```{r}
covid_df_all_states_daily_sum <- covid_df_all_states_daily %>% 
  group_by(Country_Region) %>% 
  summarise(tested = sum(daily_tested),
            positive = sum(daily_positive),
            active = sum(active),
            hospitalized = sum(hospitalizedCurr)) %>% 
  arrange(desc(tested))

covid_df_all_states_daily_sum
```

## Top 10 countries by tested persons

```{r}
covid_top_10 <- covid_df_all_states_daily_sum %>%
  head(10)

covid_top_10  
```
## Countries with highest positive cases against number of tests

```{r}
countries <- covid_top_10$Country_Region
tested_cases <- covid_top_10$tested
positive_cases <- covid_top_10$positive
active_cases <- covid_top_10$active
hospitalized_cases <- covid_top_10$hospitalized

names(tested_cases) <- countries
names(positive_cases) <- countries
names(active_cases) <- countries
names(hospitalized_cases) <- countries

positive_tested <- positive_cases / tested_cases

positive_tested_top_3 <- positive_tested[c("United Kingdom","United States", "Turkey")]

positive_tested_top_3 <- positive_tested_top_3 %>%
  round(3)
```
## Information available for this 3 countries

```{r}
united_kingdom <- c(0.11, 1473672, 166909, 0, 0)
united_states <- c(0.10, 17282363, 1877179, 0, 0)
turkey <- c(0.08, 2031192, 163941, 2980960, 0)

covid_mat <- rbind(united_kingdom, united_states, turkey)
colnames(covid_mat) <- c("Ratio", "tested", "positive", "active", "hospitalized")

covid_mat
```
## Store de information

```{r}

question <- "Which countries have had the highest number of positive cases against the number of tests?"

answer <- c("Positive tested cases" = positive_tested_top_3)

dataframes <- list(
  original = covid_df, 
  allstates = covid_df_all_states, 
  daily = covid_df_all_states_daily, 
  top_10 = covid_top_10 
  )

matrices <- list(covid_mat)
  
vectors <- list(vector_cols, countries)

data_structure_list <- list("datasets" = dataframes,
                            "matrix" = matrices,
                            "vector" = vectors)

covid_analysis_list <- list(question, answer, data_structure_list)

# solution
covid_analysis_list[[2]]

```





