---
title: "Investigating Fandango Movie Ratings"
author: "Ana Alvarez"
date: "9/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
In October 2015, a data journalist named Walt Hickey analyzed movie ratings data and found strong evidence to suggest that Fandango's rating system was biased and dishonest (Fandango is an online movie ratings aggregator). He published his analysis in [this article](https://fivethirtyeight.com/features/fandango-movies-ratings/).
* The actual rating was almost always rounded up to the nearest half-star. 
* In the case of 8% of the ratings analyzed, the rounding up was done to the nearest whole star.
* For one movie rating, the rounding off was completely bizarre: from a rating of 4 in the HTML of the page to a displayed rating of 5 stars.

In this project, we'll analyze more recent movie ratings data to determine whether there has been any change in Fandango's rating system after Hickey's analysis.

Walt Hickey made the data he analyzed publicly available on GitHub. We'll use the data he collected to analyze the characteristics of Fandango's rating system previous to his analysis:
```{r message=FALSE}
library(tidyverse)

fandango_previous <- read_csv("fandango_score_comparison.csv")
glimpse(fandango_previous)
```

One of Dataquest's team members collected movie ratings data for movies released in 2016 and 2017. The data is publicly available on GitHub and we'll use it to analyze the rating system's characteristics after Hickey's analysis.
```{r message=FALSE}
fandango_after <- read_csv("movie_ratings_16_17.csv")
glimpse(fandango_after)
```
## Exploring the data
```{r}
dim(fandango_previous)
dim(fandango_after)

```
`fandango_score_comparison.csv` contains every film that has a Rotten Tomatoes rating, a RT User rating, a Metacritic score, a Metacritic User score, and IMDb score, and at least 30 fan reviews on Fandango. The data from Fandango was pulled on Aug. 24, 2015.

```{r}
fandango_previous <- fandango_previous %>% 
  select('FILM', 'Fandango_Stars', 'Fandango_Ratingvalue', 'Fandango_votes', 'Fandango_Difference')

fandango_previous
```

Movie ratings data for 214 of the most popular movies (with a significant number of votes) released in 2016 and 2017:

```{r}
fandango_after <- fandango_after %>% 
  select('movie', 'year', 'fandango')

fandango_after
```

Initial Goal: determine whether there has been any change in Fandango's rating system after Hickey's analysis.

Because we want to find out whether the parameters of this population changed after Hickey's analysis, we're interested in sampling the population at two different periods in time — previous and after Hickey's analysis — so we can compare the two states.

The data we're working with was sampled at the moments we want: one sample was taken previous to the analysis, and the other after the analysis. We want to describe the population, so we need to make sure that the samples are representative, otherwise we should expect a large sampling error and, ultimately, wrong conclusions.

`fandango_before` only includes movies with at least 30 fan ratings on Fandango's website at the time of sampling (Aug. 24, 2015) and have had tickets on sale in 2015. So is not a random sample.

`fandango_after` collects movies released in 2016 or later and with considerable number of votes and reviews. It's unlikely to be representative of our population of interest.

Due to these limitations, our new goal is to determine whether there's any difference between Fandango's ratings for popular movies in 2015 and Fandango's ratings for popular movies in 2016.

##  Isolate only the movies released in 2015 from Hickey's dataset

```{r}
head(fandango_previous)

# Year is in the name of the film
fandango_previous <- fandango_previous %>% 
  mutate(year = str_sub(FILM, -5, -2))

```
```{r}
 # frequency distribution for the Year column

fandango_previous %>% 
  group_by(year) %>% 
  summarise(number = n())

table(fandango_previous$year)

# isolate 2015 movies

fandango_2015 <- fandango_previous %>% 
  filter(year == 2015)
```

# Isolate 2016 movies from the other dataset
```{r}
table(fandango_after$year)

fandango_2016 <- fandango_after %>% 
  filter(year == 2016)
```
# Comparing Distribution Shapes for 2015 and 2016
```{r}
fandango_2016
```


```{r}
# kernel density plots

ggplot(data = fandango_2015, 
       aes(x = Fandango_Stars)) +
  geom_density() +
  geom_density(data = fandango_2016, aes(x = fandango), color = "blue") +
  labs(title = "Comparing distribution shapes for Fandango's ratings\n(2015 vs 2016)",
       x = "Stars",
       y = "Density") +
  scale_x_continuous(breaks = seq(0, 5, by = 0.5),
                     limits = c(0, 5))
  
```
Conclussions:

* The left skew suggests that movies on Fandango are given mostly high and very high fan ratings.
* Ratings were slightly lower in 2016 compared to 2015.

```{r}
nrow(fandango_2015)
nrow(fandango_2016)
```
Datasets have different number of rows, so it doesn't make sense to compare the two tables using absolute frequencies. We can try with proportions or percentages:

```{r}
fandango_2015 %>% 
  group_by(Fandango_Stars) %>% 
  summarise(Prop_2015 = n() / nrow(fandango_2015) * 100)

fandango_2016 %>% 
  group_by(fandango) %>% 
  summarise(Prop_2016 = n() / nrow(fandango_2016) * 100)
```
Ratings of 5.0 fall from almost 7% in 2015 to less than 1% in 2016. Ratings of 4.5 fall from 38% to 25%.
The difference in ratings is clear.

```{r}
# mean, median, and mode for each distribution.

mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

stats_2015 <- fandango_2015 %>% 
  summarise(mean = mean(Fandango_Stars),
            median = median(Fandango_Stars),
            mode = mode(Fandango_Stars))


stats_2016 <- fandango_2016 %>% 
  summarise(mean = mean(fandango),
            median = median(fandango),
            mode = mode(fandango))

stats <- rbind(stats_2015, stats_2016)
stats <- stats %>% 
  mutate(year = c("2015", "2016")) %>% 
  select(year, mean, median, mode)

stats %>% 
  pivot_longer(cols = c(mean, median, mode),
               names_to = "statistics") %>% 
  ggplot(aes(x = statistics, y = value, fill = year)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = "Comparing summary statistics: 2015 vs 2016",
         y = "Stars") +
    geom_text(aes(label=round(value, 2)), position=position_dodge(width=0.9), vjust=-0.25) +
    ylim(0,5) +
    theme_classic()
```
Mean rating is 0.2 points lower in 2016 than in 2015.

There's indeed a slight difference between Fandango's ratings for popular movies in 2015 and Fandango's ratings for popular movies in 2016.

On average, popular movies released in 2016 were rated lower on Fandango than popular movies released in 2015.





