---
title: "Finding the Best Markets to Advertise In"
author: "Ana Alvarez"
date: "13/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We will analize data from freeCodeCamp's 2017 New Coder Survey (freeCodeCamp is a free e-learning platform that offers courses on web development). Our goal is to find out the two best markets to advertise in for an e-learning company that offers courses on programming.

The survey data is publicly available in [this GitHub repository](https://github.com/freeCodeCamp/2017-new-coder-survey).

We want to answer questions about a population of new coders that are interested in the subjects we teach. We'd like to know:

    * Where are these new coders located.
    * What are the locations with the greatest number of new coders.
    * How much money new coders are willing to spend on learning.

# First step: Is data representative for our population of interest?

```{r}
library(dplyr)
library(readr)

survey <- read_csv("2017-fCC-New-Coders-Survey-Data.csv")

```

Population of new coders that are interested in the subjects we teach:
```{r}
glimpse(survey)

survey %>%
  group_by(JobRoleInterest) %>%
  summarise(freq = round(n()*100/nrow(survey), 2)) %>%
  arrange(desc(freq))
```
Ranking of interests:
. Web development related roles
. Mobile development related roles
. Others

Many respondents are interested in more than one subject. Steps:
* Drop the NA values
* Split each string in the `JobRoleInterest` column to find the number of options for each participant.

```{r}
number_interests <- survey %>% 
  select(JobRoleInterest) %>% 
  tidyr::drop_na() %>% 
  rowwise() %>%  # to apply method by row and not by column
  mutate(opts = length(stringr::str_split(JobRoleInterest, ",")[[1]]))

number_interests %>% 
  ungroup() %>% # to change the rowwise
  group_by(opts) %>% 
  summarise(freq = round(n()*100 / nrow(number_interests), 2))
```
Majority of students have mixed interests. But given that we offer courses on various subjects, this can't be a problem.

```{r}
# how many respondents chose web or mobile development

survey %>% 
  mutate(web_or_mobile = stringr::str_detect(JobRoleInterest, "Web Developer|Mobile Developer")) %>% 
  tidyr::drop_na(JobRoleInterest) %>% 
  group_by(web_or_mobile) %>% 
  summarise(count = n()) %>% 
  mutate(freq = count*100 /sum(count))

```


# Where are new coders located?

origin = CountryCitizen variable
live = CountryLive variable

We'll work with the CountryLive variable because we're interested where people actually live at the moment when we run the ads.

```{r}
# respondents of job role interest
survey_good <- survey %>%
  tidyr::drop_na(JobRoleInterest) 

# group the data by country
survey_good <- survey_good %>% 
  tidyr::drop_na(CountryLive) 

survey_good %>% 
  group_by(CountryLive) %>% 
  summarise(freq = n(), percent = n()*100 / nrow(survey_good)) %>% 
  arrange(-freq)
```
We select the first four countries, because they also are english speakers.

# Money for Learning per month
```{r}
survey_good %>% 
  filter(CountryLive %in% c('United States of America', 'India', 'United Kingdom', 'Canada')) %>% 
  tidyr::drop_na(MoneyForLearning) %>% 
  group_by(CountryLive) %>% 
  summarise(money_mean = mean(MoneyForLearning)) %>% 
  arrange(-money_mean)

# Rows with MonthsProgramming == 0
survey_good %>% 
  select(MonthsProgramming) %>% 
  filter(MonthsProgramming == 0) %>% 
  nrow()

# Replace 0s with 1s to avoid division by 0
survey_good <- survey_good %>% 
  mutate(MonthsProgramming = case_when(MonthsProgramming == 0 ~1, 
                                       TRUE ~ MonthsProgramming)) %>% 
  mutate(money_per_month = MoneyForLearning / MonthsProgramming)

# Rows with money_per_month != NA
survey_good  <- survey_good %>% 
  tidyr::drop_na(money_per_month)

# Group by country
survey_good %>% 
  filter(CountryLive %in% c('United States of America', 'India', 'United Kingdom', 'Canada')) %>% 
  group_by(CountryLive) %>% 
  summarise(money_mean = round(mean(money_per_month), 1)) %>% 
  arrange(-money_mean)
```
It's surprising that people from UK spend less than people from India. 

# Removing outliers
```{r}
# Visualizing distribution of the `money_per_month` variable

only_4 <- survey_good %>% 
  filter(CountryLive %in% c('United States of America', 'India', 'United Kingdom', 'Canada'))

nrow(only_4)
nrow(survey_good)

library(ggplot2)

only_4 %>% 
  ggplot(aes(x = CountryLive, y = money_per_month)) +
  geom_boxplot() +
  ggtitle("Money Spent per Month and Country") +
  xlab("Country") +
  ylab("Money per Month") +
  theme_bw()
```
We can see values in US above \$40,000 per month. We'll remove every value that goes over \$20,000 per month.

# Removing outliers from US

```{r}
survey_good <- survey_good %>% 
  filter(money_per_month < 20000)

countries_mean = survey_good %>%
  filter(CountryLive %in% c('United States of America', 'India', 'United Kingdom', 'Canada')) %>%
  group_by(CountryLive) %>%
  summarise(mean = mean(money_per_month)) %>%
  arrange(desc(mean))

survey_good %>%
  filter(CountryLive %in% c('United States of America', 'India', 'United Kingdom', 'Canada')) %>% 
  ggplot(aes(x = CountryLive, y = money_per_month)) +
  geom_boxplot() +
  ggtitle("Money Spent Per Month Per Country\n(Distributions)") +
  xlab("Country") +
  ylab("Money per month (US dollars)") +
  theme_bw()
  
```

# Analyzing outliers

```{r}
# india outliers
survey_good %>% 
  filter(CountryLive == "India" & money_per_month > 2500) %>% 
  select(money_per_month, BootcampName, BootcampFinish, BootcampLoanYesNo, BootcampRecommend)
```
No one of the India outliers seem to have done a bootcamp. The amount looks not very realistic. We are going to remove this rows.
```{r}

# remove india outliers
survey_good <- survey_good %>% 
  filter(!(CountryLive == "India" & money_per_month > 2500))

survey_good %>%
  filter(CountryLive %in% c('United States of America', 'India', 'United Kingdom', 'Canada')) %>% 
  ggplot(aes(x = CountryLive, y = money_per_month)) +
  geom_boxplot() +
  ggtitle("Money Spent Per Month Per Country\n(Distributions)") +
  xlab("Country") +
  ylab("Money per month (US dollars)") +
  theme_bw()


```
Results look more realistic. 
```{r}
# us outliers
survey_good %>% 
  filter(CountryLive == "United States of America" & money_per_month >= 6000)

# remove outliers didn't attend Bootcamp
survey_good <- survey_good %>% 
  filter(!(CountryLive == "United States of America" & 
           money_per_month >= 6000  &
           AttendedBootcamp == 0))

# remove outliers that had been programming for less than 3 months and attend bootcamps
survey_good %>% 
  filter(CountryLive == "United States of America" & 
           money_per_month >= 6000  &
           MonthsProgramming <= 3) %>% 
  select(money_per_month, MonthsProgramming)

survey_good <- survey_good %>% 
  filter(!(CountryLive == "United States of America" & 
           money_per_month >= 6000  &
           MonthsProgramming <= 3)) 
```
Out of 11 outliers, 6 made bootcamps but 5 didn't. We remove these 5 rows. The others seem have paid a lot of money at the beginning to enroll a bootcamp and the money per month is very high.

```{r}
survey_good %>%
  filter(CountryLive %in% c('United States of America', 'India', 'United Kingdom', 'Canada')) %>% 
  ggplot(aes(x = CountryLive, y = money_per_month)) +
  geom_boxplot() +
  ggtitle("Money Spent Per Month Per Country\n(Distributions)") +
  xlab("Country") +
  ylab("Money per month (US dollars)") +
  theme_bw()
```
```{r}
# canada outlier
survey_good %>% 
  filter(CountryLive == "Canada" & money_per_month > 4500) %>% 
  select(MonthsProgramming, money_per_month)
  

```
He seems to have paid a large sum of money in the beginning to enroll in a bootcamp, and then he probably didn't spend anything for the next couple of months after the survey. We'll take the same approach here as for the US and remove this outlier.

```{r}
survey_good <- survey_good %>% 
  filter(!(CountryLive == "Canada" & money_per_month > 4500))

survey_good %>%
  filter(CountryLive %in% c('United States of America', 'India', 'United Kingdom', 'Canada')) %>% 
  ggplot(aes(x = CountryLive, y = money_per_month)) +
  geom_boxplot() +
  ggtitle("Money Spent Per Month Per Country\n(Distributions)") +
  xlab("Country") +
  ylab("Money per month (US dollars)") +
  theme_bw()
```
Let's recompute the mean values.

```{r}
survey_good %>% 
  filter(CountryLive %in% c('United States of America', 'India', 'United Kingdom', 'Canada')) %>% 
  group_by(CountryLive) %>% 
  summarise(money_mean = round(mean(money_per_month), 1)) %>% 
  arrange(-money_mean)
```
Considering the results we've found so far, one country we should definitely advertise in is the US. There are a lot of new coders living there and they are willing to pay a good amount of money each month. The second country could be Canada. But we think India is an incresing market with good perspective.





